---
title: "Review of Variable Selection"
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
```

This page contains a combination of traditional lecture materials (slides) and code demonstrating the relevant methods. The short course will proceed by working through both. We will use several recent packages in our examples; see the [About](./About.html) page for information about the package versions.

```{r setup}
library(tidyverse)
library(grpreg)
library(splines)
library(refund)
library(refund.shiny)
```


## Methods for variable selection

<script async class="speakerdeck-embed" data-id="a3d65d19df9d45f69010597ab6c957c6" data-ratio="1.33333333333333" src="//speakerdeck.com/assets/embed.js"></script>
<div style="margin-bottom:5px"> <strong> <a href="https://speakerdeck.com/jeffgoldsmith/modeling-motor-learning" title="Review of methods" target="_blank">Review of methods</a> </strong> from <strong><a href="https://speakerdeck.com/jeffgoldsmith" target="_blank">Jeff Goldsmith</a></strong>. </div>

## Practical examples

(code and text for group penalization, especially MCP)


