---
title: "Review of Variable Selection"
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
```

This page contains a combination of traditional lecture materials (slides) and code demonstrating the relevant methods. The short course will proceed by working through both. We will use several recent packages in our examples; see the [About](./About.html) page for information about the package versions.

```{r setup}
library(tidyverse)
library(grpreg)
library(splines)
library(refund)
library(refund.shiny)
```


## Methods for variable selection

<script async class="speakerdeck-embed" data-id="a3d65d19df9d45f69010597ab6c957c6" data-ratio="1.33333333333333" src="//speakerdeck.com/assets/embed.js"></script>
<div style="margin-bottom:5px"> <strong> <a href="https://speakerdeck.com/jeffgoldsmith/vs-in-fda-short-course-i" title="VS in FDA Short Course I" target="_blank">VS in FDA Short Course I</a> </strong> from <strong><a href="https://speakerdeck.com/jeffgoldsmith" target="_blank">Jeff Goldsmith</a></strong>. </div>

## Practical example

In this section we will briefly examine implementations of variable selection methods, with emphasis on group variable selection methods using penalized approaches. The primary package and function is `grpreg::grpreg`; we'll start with the example included in the help file for this function.

A brief examination of the example data is below. Particular attention should be paid to variable classes.
```{r grpreg_BWT_explore, eval = TRUE}
data(Birthwt)
names(Birthwt)

class(Birthwt$X)
class(Birthwt$group)

head(Birthwt$X) %>% round(3)
Birthwt$group
```

Next we'll fit a regression for birthweight as a response and the grouped variables as predictors.
```{r grpreg_BWT_fit, eval = TRUE}
X <- Birthwt$X
group <- Birthwt$group
y <- Birthwt$bwt

fit_lasso <- grpreg(X, y, group, penalty="grLasso")
fit_MCP <- grpreg(X, y, group, penalty="grMCP")
fit_SCAD <- grpreg(X, y, group, penalty="grSCAD")

par(mfrow = c(1,3))
plot(fit_lasso)
plot(fit_MCP)
plot(fit_SCAD)
par(mfrow = c(1,1))
```

Examining coefficients at a few choices of the tuning parameter is also informative.
```{r grpreg_coefs, eval = TRUE}
fit_MCP$beta[,c(5,10,70)]
```

The `grpreg` package has two approaches for choosing the tuning parameter: selection using an information criterion and cross validation. The code below briefly compares these approaches using a Lasso penalty.
```{r grpreg_select, eval = TRUE}
coef_BIC = grpreg::select(fit_lasso, "BIC")$beta
coef_AIC = grpreg::select(fit_lasso, "AIC")$beta
fit_cv = cv.grpreg(X, y, group, penalty = "grLasso", seed = 1)
coef_CV = fit_cv$fit$beta[,fit_cv$min]

cbind(coef_BIC, coef_AIC, coef_CV)
```


## Try it yourself!

Next we'll take a few minutes for some independent coding. Using the DTI dataset, create predictors based on averaging across profiles. Grouping these within tracts (or within modalities), use these as predictors of _pasat_ or _case_.

```{r grpreg_DTI, eval = FALSE, echo = FALSE}
load("./DataCode/DTI.RDA")

predictor_names = paste(rep(c("CSTR", "CSTL", "CCA", "OPRL", "OPRR"), each = 6), 
									rep(c("ani", "l0", "lt", "md", "mtr", "t2"), 5), 
									sep = "_")
tract_means = map(predictor_names, ~apply(get(.x), 1, mean, na.rm = TRUE))

X = matrix(unlist(tract_means), 211, 30, byrow = FALSE)
y = outcome_data$pasat

keep = complete.cases(cbind(y, X))
X = X[keep,]
y = y[keep]

group = factor(rep(c("CSTR", "CSTL", "CCA", "OPRL", "OPRR"), each = 6), 
               levels = c("CSTR", "CSTL", "CCA", "OPRL", "OPRR"))

fit_lasso <- grpreg(X, y, group, penalty="grLasso")
fit_MCP <- grpreg(X, y, group, penalty="grMCP")
fit_SCAD <- grpreg(X, y, group, penalty="grSCAD")

par(mfrow = c(1,3))
plot(fit_lasso)
plot(fit_MCP)
plot(fit_SCAD)
par(mfrow = c(1,1))

coef_BIC = grpreg::select(fit_lasso, "BIC")$beta
coef_AIC = grpreg::select(fit_lasso, "AIC")$beta
fit_cv = cv.grpreg(X, y, group, penalty = "grLasso", seed = 1)
coef_CV = fit_cv$fit$beta[,fit_cv$min]

cbind(coef_BIC, coef_AIC, coef_CV)
```
