---
title: "Other Topics"
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
```

This page contains a combination of traditional lecture materials (slides) and code demonstrating the relevant methods. The short course will proceed by working through both. We will use several recent packages in our examples; see the [About](./About.html) page for information about the package versions.

```{r setup}
library(tidyverse)
library(vbvs.concurrent)
library(refund.shiny)
```


## Other topics

<script async class="speakerdeck-embed" data-id="2db7348be17b4d47a9b0aca62b482559" data-ratio="1.33333333333333" src="//speakerdeck.com/assets/embed.js"></script>
<div style="margin-bottom:5px"> <strong> <a href="https://speakerdeck.com/jeffgoldsmith/vs-in-fda-short-course-iv" title="VS in FDA Short Course IV" target="_blank">VS in FDA Short Course IV</a> </strong> from <strong><a href="https://speakerdeck.com/jeffgoldsmith" target="_blank">Jeff Goldsmith</a></strong>. </div>


## Practical example

Due to a lack of publicly available and easy-to-implement code for scalar-on-image regression, we focus on methods for the functional linear concurrent model. 

The snippet below simulates a simple example which we will use for illustration. 

```{r vbvs_sim_data}
## set design elements
set.seed(1)
I = 50

## coefficient functions
beta1 = function(t) { 1 }
beta2 = function(t) { cos(2*t*pi) }
psi1 = function(t) { sin(2*t*pi) }

## generate subjects and observation times
concurrent_data = 
  data.frame(
    subj = rep(1:I, each = 20)
  ) %>%
  mutate(time = runif(dim(.)[1])) %>%
  arrange(subj, time) %>%
  group_by(subj) %>%
    mutate(Cov_1 = runif(1, .5, 1.5) * sin(2 * pi * time),
           Cov_2 = runif(1, 0, 1) + runif(1, -.5, 2) * time,
    			 FPC_score = rnorm(1, 0, .5),
           Y = Cov_1 * beta1(time) + 
               Cov_2 * beta2(time) +
    			 	   FPC_score * psi1(time) + 
               rnorm(20, 0, .5)) %>%
  ungroup() %>%
	dplyr::select(subj, time, Y, everything())

superfulous_covariates = matrix(rnorm(18 * I * 20),	nrow = I * 20, ncol = 18)
colnames(superfulous_covariates) = paste0("Cov_", 3:20)

concurrent_data = bind_cols(concurrent_data, as.data.frame(superfulous_covariates))
```

The plot below shows the first two predictors and the response, highlighting four subjects.

```{r plot_data, fig.align='center', fig.height=3, fig.width=9, echo=FALSE}
library(gridExtra)

plot_function = function(value_var, data) {
  ggplot(data, aes_string(x = "time", y = value_var, group = "subj")) + geom_path(alpha = .1) +
    theme_bw() +
    geom_path(data = filter(data, subj %in% 1:4) %>% mutate(subj = as.factor(subj)), aes(color = subj)) + 
    theme(legend.position = "none")
}

panels = lapply(c("Cov_1", "Cov_2", "Y"), plot_function, data = concurrent_data)

grid.arrange(panels[[1]], panels[[2]], panels[[3]], nrow = 1)

```

To fit the functional linear concurrent model with variable selection, we can use `vbvs_concurrent`. 

```{r vbvs_fit}
fit_vbvs = vbvs_concurrent(Y ~ Cov_1 + Cov_2 + Cov_3 + Cov_4 + Cov_5 + 
													 Cov_6 + Cov_7 + Cov_8 + Cov_9 + Cov_10 + 
													 Cov_11 + Cov_12 + Cov_13 + Cov_14 + Cov_15 + 
													 Cov_16 + Cov_17 + Cov_18 + Cov_19 + Cov_20 | time, 
													 id.var = "subj", data = concurrent_data, 
                           t.min = 0, t.max = 1, standardized = TRUE)
```

The plot below shows true coefficients and estimates without using variable selection.

```{r coefficients, echo=FALSE, fig.height=6, fig.width=9, echo=FALSE}
coefs = coef(fit_vbvs, t.new = seq(0, 1, length = 101)) %>%
  mutate(True_1 = beta1(time),
         True_2 = beta2(time)) %>%
  gather(key, value, Cov_1:True_2) %>%
  extract(key, c("Estimate", "Beta"), "([[:alnum:]]+)_([[:alnum:]]+)") %>%
	mutate(Beta = factor(Beta, levels = 1:20))

ggplot(coefs, aes(x = time, y = value, color = Estimate)) + geom_path() + 
  facet_wrap(~Beta) + theme_bw() + theme(legend.position = "bottom")
```

Interactive graphics show observed data, coefficient functions, and residual curves. The code below will produce such a graphic. As noted in the slides, this approach also uses FPCA to decompose residual curves as part of the overall model fitting; this decomposition can be viewed using the tool we saw previously.

```{r interactive_plot, eval=FALSE}
plot_shiny(fit_vbvs)
plot_shiny(fit_vbvs$fpca.obj)
```


## Try it yourself!

As in all other examples, the preceding approach has tuning parameters. These can be set by hand using the _v0_ argument to the `vbvs_concurrent` function. There is also a `cv_vbvs_concurrent` function, and versions without variable selection.

As an alternative to the functional linear concurrent model, note that the functional linear model is a special case; by carefully constructing the data frame containing predictions, `vbvs_concurrent` can be used for variable selection in that setting, and is interesting to compare to other methods we've seen so far.
