---
title: "Function-on-Scalar Regression"
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
```

This page contains a combination of traditional lecture materials (slides) and code demonstrating the relevant methods. The short course will proceed by working through both. We will use several recent packages in our examples; see the [About](./About.html) page for information about the package versions.

```{r setup}
library(tidyverse)
library(grpreg)
library(splines)
library(refund)
library(refund.shiny)
```

In this section we'll use the [Activity data](./Data.html) to motivate the function-on-scalar regression model and variable selection in this context. Our main outcome of interest is the observed physical activity profile, with predictors including age, sex, and season. Below we plot activity profiles colored by sex.


## Methods for variable selection

Slides below review function-on-scalar regression and introduce the ideas for variable selection in this context.

(placeholder for slides: review fosr using splines; variable selection using group penalties; correlated errors (GLS or Bayes methods); discuss challenges, including sparse or incomplete data, smoothness constraints, and scaling)

## Practical example

(code and text for OLS-based fosr in the accelerometer example; include Yakuan's function and results)


## Try it yourself!

Using the DTI data with several predictors. 


